---
title: "There's a tool to do that"
author: "Russ Hyde"
date: "14 February 2019"
output:
  ioslides_presentation:
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

<!-- Formatting code
     ---------------
     Thanks to Yihui for text-display code:
    - https://stackoverflow.com/questions/29257985/r-markdown-how-do-i-show-file-contents

    Thanks to @hrbrmstr for the scrolling text-box styling:
    - https://stackoverflow.com/questions/38578152/make-scrollbar-appear-in-rmarkdown-code-chunks-html-view
-->

```{css}
pre code, pre, code {
  max-height: 550px;
  white-space: pre !important;
  overflow-x: scroll !important;
  overflow-y: scroll !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
}
```

## R / Python development tools

```{r}
# library(kableExtra)
library(knitr)
library(magrittr)
library(tibble)
```

```{r, echo  FALSE}
tribble(
  ~ Job, ~ Python, ~ R,
  "IDE",  "PyCharm",  "Rstudio",
  "Styling", "black", "styler",
  "Linting", "pylint", "lintr",
  "Testing", "pytest", "testthat",
  "Packages", "cookiecutter / setuptools", "devtools / usethis",
  "Docs", "sphinx / restructuredText", "roxygen2 / rmarkdown"
  ) %>%
  kable()
```

<!-- Give more detail / alternatives in specific sections -->
<!-- Other jobs:
    "Testing (Advanced)": "hypothesis / raygun", "hedgehog"
    "CI": Travis
    "Release": CRAN / Anaconda / pip / pypi
    "VCS": git / github
-->

<!-- Mention:
    - Style guides for code and internal consistency
    - Automatic documentation generation
    - Code review  / Stack overflow
    - Beyond code-style:
        - goodpractice
        - dupree / pylint's duplicate finder
        - mypy & typing
    - Awesome static analysis
    - CRAN packageDev Task View
    - Dangers of `fixing` poorly-tested code
    - Performance profiling
    - Lots of python 2 legacy code is going to be chucked off a cliff
      - 2to3.py
-->



# Rosalind

## Initial Solution for `Longest Increasing Subsequence` from rosalind.info

[Rosalind](http://rosalind.info/about/)

[Longest Increasing Subsequence](http://rosalind.info/problems/lgis/)

##

```{bash, comment = ''}
cat rosa/initial.py
```

# Automate reviewing code-quality / consistency: `pylint`

## `pylint` messages

<!--
    [From realpython.com/python-code-quality]

    Pylint prefixes each of the problem areas with a R, C, W, E, or F, meaning:

    [R]efactor for a “good practice” metric violation
    [C]onvention for coding standard violation
    [W]arning for stylistic problems, or minor programming issues
    [E]rror for important programming issues (i.e. most probably bug)
    [F]atal for errors which prevented further processing
-->

```{bash, echo = TRUE}
echo ...
pylint --long-help | tail -n20 | head -n 8
```

## `pylint` return codes

```{bash, echo = TRUE}
echo ...
pylint --long-help | tail -n12
```

## Convention (C) and Warning (W) lints in the initial code

<!--
    Have to use --exit-zero since non-zero bash blocks kill `Rmarkdown`
    compilation
-->

```{bash, echo = FALSE}
echo "" > rosa/cleaned.py
pylint rosa/cleaned.py --exit-zero > /dev/null
```

```{bash, eval = TRUE, echo = TRUE}
cp rosa/{initial,cleaned}.py

pylint rosa/cleaned.py --exit-zero
```

# Automate Styling : `black`

## `black` modifies files in-place

```{bash, echo = TRUE}
black rosa/cleaned.py
```

## Example `black` differences

```{bash, echo = TRUE}
diff rosa/initial.py rosa/cleaned.py | grep -e ">\|<" | head -n6
```

## The styled code

```{bash, comment = ''}
cat rosa/cleaned.py
```

## `black` reduced the lint count

```{bash, echo = TRUE}
pylint rosa/cleaned.py --exit-zero
```

## Manually fixed the remaining lints:

Changes:

- Added module / function description strings

- Varnames:
    - Global constant vars $\rightarrow$ upper-case
    - Short varnames fixed `lt` $\rightarrow$ `parents`

- Removed an unused argument

- Changed empty-list tests:
    - `if len(my_list) > 0: ...` $\rightarrow$ `if my_list: ...`

## Styled & Lint-free code:

```{bash, comment = ''}
cat rosa/nolints.py
```

## Passing lints != Good code

- [extensibility / maintainability]
    - Untestable / unimportable
    - Pulls data from a hard-coded filepath

- [correctness]
    - edge-case failures (see later)

- [idioms]
    - Use `with` when opening files
    - Poor loops: `for ... in range(...)`, `i in my_dict.keys()`

- [general]
    - Poor docs / Poor names / Nested-conditionals / Long-function

[Python Book of Antipatterns](
  https://docs.quantifiedcode.com/python-anti-patterns/
)

## Passing lints != Importable code

```{python, echo = FALSE}
import os, sys

sys.path.insert(0, os.getcwd())
```

```{python, echo = TRUE}
try:
    import rosa.nolints
except FileNotFoundError as e:
    print(e)
```

# Automate running your tests: `pytest`

## Make it importable / testable

We put a guard-block around the global env code

```{bash, echo = TRUE}
cat rosa/testable.py | grep -A7 "__main__"
```

```{python, echo = TRUE}
from rosa.testable import lgis # successful import
```

## Fix up with tests

The function works fine, though, doesn't it ....

```{python, echo = TRUE}
print(
  lgis([1, 3, 4,
           2,
                 5, 7,
              4])
)
```

## We can write tests to check the function

```{python, echo = TRUE}
# I've also put this code into tests/test_lgis.py

from rosa.testable import lgis

def test_it_isnt_greedy():
    assert lgis([1, 4, 2, 3, 5]) != [1, 4, 5]

def test_increasing_sequence():
    assert lgis([1, 3, 4, 2, 5, 7, 4]) == [1, 3, 4, 5, 7]
```

```{python, echo = TRUE}
test_it_isnt_greedy()
test_increasing_sequence()

# silent means they've passed
```

## `pytest` requires a python package structure

- `__setup__.py` in the main directory with (at least) this code:

```{bash, echo = TRUE}
echo -e \
"from setuptools import setup, find_packages
setup(name='rosa', packages=find_packages())" \
  > setup.py
```

- `__init__.py` in every dir containing your source code

```{bash, echo = TRUE}
echo "name = 'rosa'" > rosa/__init__.py
```

## Install the package and run the tests

```{bash, echo = TRUE, results = FALSE}
pip install -e .
```

```{bash, echo = TRUE}
pytest
```

# Automate generating test cases: `hypothesis`

## Randomly create your test data

[hypothesis.readthedocs.io](https://hypothesis.readthedocs.io/en/latest/)

```{python, echo = TRUE}
from hypothesis import given, strategies as st
```

```{python, echo = TRUE}
@given(my_ints=st.lists(st.integers(min_value=1), max_size=1))
def test_trivial_sequences(my_ints):
    assert my_ints == lgis(my_ints)
```

This generates

- lots of separate integer-lists (`st.lists(st.integers, ....)`)

- each list contains at-most one element (`... max_size=1`)

- each integer is $\geq1$ (`st.integers(min_value=1)`)

## But ...

```{python, echo = TRUE}
try:
    test_trivial_sequences()
except Exception as e:
    print(e)
```

Randomised tests identified an edge-case failure for `lgis`: empty-list

Any data that fails should be converted into a unit test

## Try again with an updated function

```{bash, comment = '', echo = TRUE}
cat rosa/fixed_for_empty_list.py | grep -A6 "def lgis"
```

```{python, echo = TRUE}
from rosa.fixed_for_empty_list import lgis

test_trivial_sequences()
```

## Will this test pass for all input?

```{python, echo = TRUE}
@given(
    my_ints=st.lists(
        st.integers(min_value=1), min_size=2, max_size=5
    )
)
def test_with_sorted_input(my_ints):
    assert sorted(my_ints) == lgis(sorted(my_ints))
```

## ... No!

```{python, echo = TRUE}
try:
    test_with_sorted_input()
except Exception as e:
    print(e)
```

This was a human error

Moral: You still have to think about the tests and the test-data

## Fix: ensure there are no repeated numbers in the input

```{python, echo = TRUE}
@given(
    my_ints=st.lists(
        st.integers(min_value=1), min_size=2, max_size=5,
        unique=True
    )
)
def test_with_sorted_unique_input(my_ints):
    assert sorted(my_ints) == lgis(sorted(my_ints))
```

```{python, echo = TRUE}
test_with_sorted_unique_input()
```

## Use-cases for randomised testing

- Identify edge-case failures that you hadn't considered

- Compare a _fancy_ algorithm to a brute-force solution (for small input)

- Check the speed of an algorithm

It forces you to think about the algebra of your problem

For the example here:

- the output should be increasing

- the output should be a subsequence of the input

- If we dovetail two sequences, the output should be no shorter than that for
either sequence