---
title: "There's a tool for that"
author: "Russ Hyde"
date: "14 February 2019"
output:
  ioslides_presentation:
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, message = FALSE}
library(knitr)
library(magrittr)
library(tibble)
library(stackr) # "dgrtwo/stackr"
library(wordcloud)
library(RColorBrewer)
```
```{r}
set.seed(1)
```

<!-- Formatting code
     ---------------
     Thanks to Yihui for text-display code:
    - https://stackoverflow.com/questions/29257985/r-markdown-how-do-i-show-file-contents

    Thanks to @hrbrmstr for the scrolling text-box styling:
    - https://stackoverflow.com/questions/38578152/make-scrollbar-appear-in-rmarkdown-code-chunks-html-view
-->

```{css}
pre code, pre, code {
  max-height: 480px;
  white-space: pre !important;
  overflow-x: scroll !important;
  overflow-y: scroll !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
}
```

## This talk

Assume some basic python (mainly; and a bit of R)

- No: _solving problems_ with code

- Yes: easing _the problems of coding_

Links

- [http://rpubs.com/russH/467834]()

- [https://github.com/russHyde/bfx_201903]()

## My expertise wordcloud (according to Stack-Overflow) { .smaller }

```{r, fig.height=3.2, warning=FALSE}
with(
  stackr::stack_users("1845650", "top-tags"),
  wordcloud::wordcloud(
    tag_name, answer_score, min.freq = 0, colors = RColorBrewer::brewer.pal(6, "Purples")[-1],
    scale = c(10, 0.5)
    ))
```

... happy to answer `R` queries through `Yammer` / email

- Made with `github::dgrtwo/stackr` and `wordcloud` packages

## What tools / libraries help you ... { .smaller }

... <span style="color:purple">package</span> stuff

- [py] cookiecutter / setuptools; [R] devtools / usethis

... <span style="color:purple">document</span> stuff

- [py] sphinx / restructuredText; [R] roxygen2

... <span style="color:purple">test</span> stuff

- [py] pytest / hypothesis; [R] testthat / hedgehog

... <span style="color:purple">style</span> stuff

- [py] black / pylint / flake8; [R] styler / lintr

... <span style="color:purple">interact</span> with the user

- [py] click; [R] optparse


<!-- Give more detail / alternatives in specific sections -->
<!-- Other jobs:
    "Testing (Advanced)": "hypothesis / raygun", "hedgehog"
    "CI": Travis
    "Release": CRAN / Anaconda / pip / pypi
    "VCS": git / github
-->

<!-- Mention:
    - Style guides for code and internal consistency
    - Automatic documentation generation
    - Code review  / Stack overflow
    - Beyond code-style:
        - goodpractice
        - dupree / pylint's duplicate finder
        - mypy & typing
    - Awesome static analysis
    - CRAN packageDev Task View
    - Dangers of `fixing` poorly-tested code
    - Performance profiling
    - Lots of python 2 legacy code is going to be chucked off a cliff
      - 2to3.py
-->



# Rosalind

## Initial Solution for `Longest Increasing Subsequence` from rosalind.info

[Rosalind](http://rosalind.info/about/)

- Bioinformatics algorithmic challenges (akin to Project Euler / Codewars /
Codesignal)

[Longest Increasing Subsequence](http://rosalind.info/problems/lgis/)

- Pulled out my solution code (written ~ 6 years ago)

- I _know_ that my code is bad. That's why I chose it.

## {.smaller}

```{r, echo = FALSE, message = FALSE}
library(tidygraph)
library(ggraph)

set.seed(0)

nodes <- data.frame(
    name = 1:9,
    value = sample(1:100, size = 9),
    label = c("start", rep(NA, 7), "end"),
    level = NA
  )

for (i in 1:9){
  parents <- which(nodes[1:i, "value"] < nodes[i, "value"])
  if (length(parents) == 0){
    nodes[i, "level"] = 1
  } else {
    nodes[i, "level"] <- 1 + max(nodes[parents, "level"])
  }
}

lgis_example <- tbl_graph(
  nodes = nodes,
  edges = data.frame(from = 1:8, to = 2:9)
) %>%
  mutate(x = name, y = value)
```

```{r, message = FALSE, warning = FALSE}
lgis_example %>%
  ggraph() +
  geom_node_point(aes(colour = as.factor(level)), size = 4) +
  geom_node_text(aes(label = label), nudge_y = -5) +
  geom_node_text(aes(label = value), nudge_y = 5, nudge_x = 0.1) +
  geom_edge_link2() +
  theme_graph() +
  scale_color_discrete(name = "Level") +
  ggtitle(
    "`LGIS` Illustration",
    subtitle = "Made in `R` using `tidygraph` / `ggraph`"
  )
```

- Input: unique positive integers

- Partition entries into levels (based on values of preceding entries)

- Identify parent(s) of each entry

## Initial Code

```{bash, echo = TRUE, comment = ''}
cat rosa/initial.py
```

# Automate packaging: `cookiecutter`

## Python packages look like this

```{bash, echo = TRUE}
tree minimal_py_package
```

- Stuff works better when you structure it better

## `cookiecutter`

Templates that help structure a programming project

- [`cookiecutter` docs](https://cookiecutter.readthedocs.io/en/latest/readme.html)

Examples:

- [Minimal python package template](
    https://github.com/kragniz/cookiecutter-pypackage-minimal)

- [`cookiecutter` author's python package template](
    https://github.com/audreyr/cookiecutter-pypackage)

- [Data science template](
    https://github.com/drivendata/cookiecutter-data-science)

- [My snakemake / conda / .Rmd analysis-project template](
    https://github.com/russHyde/data_buddy)
    - work in progress

## Define a minimal python package {.smaller}

To use a `cookiecutter`:

```
cookiecutter git@github.com:<some_repo>.git
```

Then answer some config questions:

```
cookiecutter git@github.com:kragniz/cookiecutter-pypackage-minimal.git

You've downloaded /home/ah327h/.cookiecutters/cookiecutter-pypackage-minimal before. Is it okay to delete and re-download it? [yes]:
author_name [Louis Taylor]: Russell Hyde
author_email [louis@kragniz.eu]: me AT somewhere.uk
package_name [cookiecutter_pypackage_minimal]: rosa_initial
package_version [0.1.0]:
...
```

## Resulting directory:

I copied in my `lgis.py` code into the resulting directory

```{bash}
tree rosa_initial
```

# Automate (trivial bits of) code-review: `pylint`

## `pylint` {.smaller}

[Link](https://www.pylint.org/)

- finds non-idiomatic and inconsistently styled bits of code

- highly configurable (alternative: `flake8`)

```{bash, echo = TRUE}
pylint --long-help | tail -n20 | head -n 8
```

## Booooring lints in the initial code {.smaller}

<!--
    Have to use --exit-zero since non-zero bash blocks kill `Rmarkdown`
    compilation
-->

```{bash, echo = TRUE}
(cd rosa_initial/ && pylint rosa_initial --exit-zero)
```

# Automate Styling : `black`

## `black` {.smaller}

[Link](https://black.readthedocs.io/en/stable/)

- formats files in-place

- uncompromising and unconfigurable

```{bash, echo = TRUE, eval = TRUE}
# - Made another package `rosa_cleaned` (not shown)
#
# - Ensure the original code is present in the `rosa_cleaned` package
# to begin with
cp rosa_initial/rosa_initial/lgis.py \
    rosa_cleaned/rosa_cleaned/lgis.py
```

```{bash, echo = TRUE}
black rosa_cleaned/*/lgis.py
```

## Example `black` differences

```{bash, echo = TRUE}
diff rosa_{initial,cleaned}/*/lgis.py | grep -e ">\|<" | head -n6
```

- These are trivial differences

- Value comes when applied to large projects

## The styled code

```{bash, comment = '', echo = TRUE}
cat rosa_cleaned/*/lgis.py
```

## After `black` the lints are non-trivial

```{bash, echo = TRUE}
(cd rosa_cleaned && pylint */lgis.py --exit-zero)
```

## Manually fixed the remaining lints:

Changes:

- Added module / function description strings

- Varnames:
    - Global constant vars $\rightarrow$ upper-case
    - Short varnames fixed `lt` $\rightarrow$ `parents`

- Removed an unused argument

- Changed empty-list tests:
    - `if len(my_list) > 0: ...` $\rightarrow$ `if my_list: ...`

## Styled & Lint-free code: {.smaller}

```{bash, comment = '', echo = TRUE}
cat rosa/nolints.py
```

## Passing lints != Good code {.smaller}

- <span style="color:purple">extensibility / maintainability</span>
    - Untestable / unimportable
    - Pulls data from a hard-coded filepath

- <span style="color:purple">correctness</span>
    - edge-case failures (see later)

- <span style="color:purple">idioms</span>
    - Use `with` when opening files
    - Poor loops: `for ... in range(...)`, `i in my_dict.keys()`

- <span style="color:purple">general</span>
    - Poor docs / Poor names / Nested-conditionals / Long-function

- <span style="color:purple">comments?</span>

[Python Book of Antipatterns](
  https://docs.quantifiedcode.com/python-anti-patterns/
)

## Passing lints != Importable code

```{python, echo = FALSE}
import os, sys

sys.path.insert(0, os.getcwd())
```

```{python, echo = TRUE}
# The lint-free code is in ./rosa/nolints.py
try:
    import rosa.nolints
except FileNotFoundError as e:
    print(e)
```

```{bash, echo = TRUE, eval = TRUE}
# Hardcoded file in global-env at the offending line:
grep "data_LGIS.txt" rosa/nolints.py
```

# Automate running your tests: `pytest`

## Make the file importable / testable

Made another python package: `rosa_testable`

- `lgis.py` now contains a guard-block around the global-env code

```{bash, echo = TRUE, comment=""}
cat rosa_testable/*/lgis.py | grep -A7 "__main__"
```

## Make the function importable / testable {.smaller}

Also, we ensured the function was exported by the `rosa_testable` package:

```{bash, echo = TRUE, comment=""}
cat rosa_testable/rosa_testable/__init__.py
```

## Check importability {.smaller}

```{bash, echo = TRUE, comment="", message=FALSE}
pip install -e rosa_testable # [bash]
```

```{python, echo = TRUE, eval = TRUE}
import rosa_testable as rt # [python] successful import
print(rt.lgis([1,4,3,6,7,2,5]))
```

## Preliminary tests {.smaller}

```{bash, comment = "", echo = TRUE}
cat rosa_testable/tests/test_lgis.py
```

```{bash, echo = TRUE, eval = TRUE, comment=""}
pytest rosa_testable --quiet # run all tests in rosa_testable/tests
```

# Automate test-case generation: `hypothesis`

## `hypothesis` {.smaller}

[Link](https://hypothesis.readthedocs.io/en/latest/)

- Randomly create your test data

```{python, echo = TRUE}
from hypothesis import given, strategies as st
```

```{python, echo = TRUE}
@given(
    my_ints=st.lists(
        st.integers(min_value=1),
        max_size=1)
    )
def test_trivial_sequences(my_ints):
    assert my_ints == rt.lgis(my_ints)
```

This generates

- lots of separate integer-lists (`st.lists(st.integers, ....)`)

- each list contains at-most one element (`... max_size=1`)

- each integer is $\geq1$ (`st.integers(min_value=1)`)

## But ...

```{python, echo = TRUE}
try:
    test_trivial_sequences()
except Exception as e:
    print(e)
```

Randomised tests identified an edge-case failure for `lgis`: empty-list

Any data that fails should be converted into a unit test

## Will this test pass for all input?

```{python, echo = TRUE}
@given(
    my_ints=st.lists(
        st.integers(min_value=1),
        min_size=2,
        max_size=5
    )
)
def test_with_sorted_input(my_ints):
    assert sorted(my_ints) == rt.lgis(sorted(my_ints))
```

## ... No!

```{python, echo = TRUE}
try:
    test_with_sorted_input()
except Exception as e:
    print(e)
```

This was a human error

Moral: You still have to think about the tests and the test-data

## Fix: ensure there are no repeated numbers in the input

```{python, echo = TRUE}
@given(
    my_ints=st.lists(
        st.integers(min_value=1),
        min_size=2,
        max_size=5,
        unique=True
    )
)
def test_with_sorted_unique_input(my_ints):
    assert sorted(my_ints) == rt.lgis(sorted(my_ints))
```

```{python, echo = TRUE}
test_with_sorted_unique_input() # the test now passes
```

## Use-cases for randomised testing

- Identify edge-case failures that you hadn't considered

- Compare a _fancy_ algorithm to a brute-force solution (for small input)

- Check the speed of an algorithm

- Provides alternative, property-based, viewpoint, eg here:
    - the output should be increasing

    - the output should be a subsequence of the input

    - If we dovetail two sequences, the output should be no shorter than that
    for either sequence

# Automate documentation generation: `sphinx` / `click`

## Documentation

- The tools above improve _your_ efficiency

- Good docs improve the effiency of the user and your collaborators (incl,
  _future you_)

- Many levels:

    - Code comments & Exception / error handling

    - [`click` / `argparse`]
        - `--help` / usage strings

    - [`sphinx` / `ReadTheDocs`]
        - API / program reference
        - Tutorials / Cheatsheets / Cookbooks etc

## Docstrings $\rightarrow$ Sphinx $\rightarrow$ Documentation

Google-formatted docstrings used here.
[Alternative formats.](https://www.datacamp.com/community/tutorials/docstrings-python)

```{bash, echo = TRUE, comment = ""}
grep -A10 "def lgis" rosa_testable/*/lgis.py
```

## Sphinx setup :0( {.smaller}

The workflow for generating docs with sphinx looks like:

```
# 1) - Setup a package structure
# 2) - Add some source code, with docstrings associated with the
# modules / classes / functions

# 3) Initialise the `docs` file structure
cd my_pkg
mkdir docs && cd docs
sphinx-quickstart # answer lots of questions
sphinx-apidoc -o source ../my_pkg
make html

# 4) Modify source/conf.py, swear a lot, question your sanity and try
# again

# 5) Ask Russ "What happened to all the automated tools you were going
# to talk about?"
```

[sphinx-quickstart options](
    https://pythonhosted.org/an_example_pypi_project/sphinx.html)

[Sam Nicholls on sphinx setup](
    https://samnicholls.net/2016/06/15/how-to-sphinx-readthedocs/)

## Try a full-featured python-package cookie-cutter

```
cookiecutter git@github.com:audreyr/cookiecutter-pypackage.git
```

This has a Makefile for: linting, testing, installation and <span style="color:purple">documentation</span>.

```
full_name [Audrey Roy Greenfeld]: Russell Hyde
email [audreyr@example.com]: russ AT somewhere.net
github_username [audreyr]: russHyde
project_name [Python Boilerplate]: detailed_py_package
project_slug [detailed_py_package]:
project_short_description [Python Boilerplate contains all the boilerplate you need to create a Python package.]: -
pypi_username [russHyde]: -
version [0.1.0]:
use_pytest [n]: y
...
```

## Detailed package structure {.smaller}

```{bash, echo = TRUE, comment = ""}
tree detailed_py_package
```

## Made a `sphinx`-ready package

```{bash, echo = TRUE}
# use `make install` in normal use
pip install -e rosa_with_docs
```

## Make the docs

```{bash, echo = TRUE, eval = FALSE}
# Not ran while building presentation
(cd rosa_with_docs && make docs)
```

[(Not on rpubs) Link to documentation](
  ./rosa_with_docs/docs/_build/html/index.html
)

```
rm -f docs/rosa_with_docs.rst
rm -f docs/modules.rst
sphinx-apidoc -o docs/ rosa_with_docs
Creating file docs/rosa_with_docs.rst.
Creating file docs/modules.rst.
make -C docs clean
make[1]: Entering directory '/home/ah327h/workshops/bfx_201903/rosa_with_docs/docs'
Removing everything under '_build'...
make[1]: Leaving directory '/home/ah327h/workshops/bfx_201903/rosa_with_docs/docs'
make -C docs html
make[1]: Entering directory '/home/ah327h/workshops/bfx_201903/rosa_with_docs/docs'
Running Sphinx v1.8.4
making output directory...
building [mo]: targets for 0 po files that are out of date
building [html]: targets for 9 source files that are out of date
updating environment: 9 added, 0 changed, 0 removed
reading sources... [100%] usage
/home/ah327h/workshops/bfx_201903/rosa_with_docs/docs/index.rst:2: WARNING: Title underline too short.

Welcome to rosa_with_docs's documentation!
======================================
looking for now-outdated files... none found
pickling environment... done
checking consistency... done
preparing documents... done
writing output... [100%] usage
generating indices... genindex py-modindex
highlighting module code... [100%] rosa_with_docs.lgis
writing additional pages... search
copying static files... WARNING: html_static_path entry '/home/ah327h/workshops/bfx_201903/rosa_with_docs/docs/_static' does not exist
done
copying extra files... done
dumping search index in English (code: en) ... done
dumping object inventory... done
build succeeded, 2 warnings.

The HTML pages are in _build/html.
make[1]: Leaving directory '/home/ah327h/workshops/bfx_201903/rosa_with_docs/docs'
python -c "$BROWSER_PYSCRIPT" docs/_build/html/index.html
```

## Easy `--help` strings / arg-parsing: `click`

Another use of docstrings

Command line programs:

- Standard solution is [`argparse`](https://docs.python.org/3/library/argparse.html)

- [`click`](http://click.palletsprojects.com/en/5.x/)
    - is less verbose
    - decorators define commands / arguments
    - docstrings define help messages

```
# Wanted
<program_name> <command_name> [arguments]
```

## Implementation with `click` {.smaller}

```{bash, echo = FALSE, eval = TRUE}
pip install -e rosa_cli 1> /dev/null
```

```{bash, echo = TRUE, comment = ""}
cat rosa_cli/rosa_cli/cli.py # I have renamed the `lgis` function
```

## Main program help-string

```{bash, echo = TRUE, comment = ""}
rosa_cli --help
```

## Subcommand help-string

```{bash, echo = TRUE, comment = ""}
rosa_cli lgis --help
```

## Run the tool

```{bash, echo = TRUE, comment = ""}
echo 6 > temp.txt
echo "1 5 3 2 6 5" >> temp.txt

rosa_cli lgis temp.txt
```

<!--
Made another project from the pypackage cookiecutter template: rosa_cli

Copied lgis and test_lgis into rosa_cli
- ensure test_lgis imports from rosa_cli
- ensure rosa_cli/__init__.py imports rosa_cli.lgis.lgis as lgis

Installed using `pip install -e .`
Tested using `make test`

Added empty-sequence test for lgis

Added randomised tests: trivial sequences, sorted-unique tests

Renamed lgis as longest_increasing_subsequence

Added function longest_decreasing_subsequence

Removed MAXNUM from __main__ guard block (since unused)

Copied code from __main__ guard block of lgis.py into cli.py

Imported longest_[in|de]creasing_subsequence into cli.py

Added `lgis` as a subcommand of the `rosa_cli` CLI application
-->

## Comparison to R tools

```{r, echo  FALSE}
tribble(
  ~ Job, ~ Python, ~ R,
  "IDE",  "PyCharm",  "Rstudio",
  "Styling", "black", "styler",
  "Linting", "pylint / flake8", "lintr",
  "Testing", "pytest / hypothesis", "testthat / hedgehog",
  "Packaging", "cookiecutter / setuptools", "devtools / usethis",
  "Docs", "sphinx / restructuredText", "roxygen2 / rmarkdown",
  "CLI", "click / argparse", "optparse"
  ) %>%
  kable()
```